---
title: 'Homework #7'
author: "Darien Lozon"
date: "April 2, 2019"
output: html_document
---

1. Load data and place into an unmarkedFramePCount object.

```{r}
# Set the stage #
setwd("C:/Users/dnl0009/Google Drive/SP19 Courses/WMAN633/")
library(unmarked)

# Load and transform data #
dat <- read.csv('count.csv')
datmat <- as.matrix(dat)

# Load covariate data # 
p <- read.csv('obs_covs.csv')
psi <- read.csv('site_covs.csv')
site_covs <- list(
   x1 = data.frame(psi[,'x1']),
   x2 = data.frame(x2 = factor(psi$x2))
)


# Set up frame #
nmix_dat <- unmarkedFramePCount(y=datmat,
                                siteCovs=psi,
                                obsCovs=list(j=p))
```

---

2. Fit an N-mixture model that assumes conditional detection probability is a function of the detection covariate provided, and expected abundance is an additive function of variables x1 and x2.

```{r}
fit <- pcount(~j~x1+x2,nmix_dat,K=100)
```

---

3. Interpret the effect of x1 on the expected count at each site. Verify your interpretation in R.

With each unit increase in x1, there is an expected log proportional change of 0.37. To verify, I will assume x2 = A and observe what happens when there is a 1-unit change in x1 (from 0 to 1). The trick is to keep in mind that abundance is Poisson, so log(lambda) = b0+b1+...+ etc., so exponentiation is required to get at the true expected count.

```{r}
# Look over values #
summary(fit)

# Verify interpretation #
b <- coef(fit) # Store coefficients for later; we'll need b[2] to compare at the end.
z <- exp(b[1])
o <- exp(b[1] + b[2])
log(o/z) # larger number over smaller
b[2]

# Another way to confirm #
o/z
exp(b[2])
```

---

4. Predict and plot the effect of the supplied detection covariate. Do this over the range of this covariate.

```{r}
new_p <- data.frame(j=seq(from=min(p),to=max(p),length.out=200))
prd <- predict(object=fit,newdata=new_p,type='det')

plot(x=new_p$j,y=exp(prd$Predicted),
     main='Effects of Detection Covariate on Estimated Abundance',
     xlab='Detection',ylab='Estimated Abundance',type='l')
lines(y=exp(prd$upper),x=new_p$j,col="red",lty=3) # Plot 95% confidence intervals
lines(y=exp(prd$lower),x=new_p$j,col="blue",lty=3)
legend("topright",c('Prediction','Upper CI','Lower CI'),col=c('black','red','blue'),lty=c(1,3,3))
```

---

5. Use contrasts to compare expected abundance between all pairwise levels of variable x2. Obtain p-values associated with each contrast and tell me whether you reject or fail to reject each null hypothesis tested.

```{r}
X <- matrix(c(0,0,1,-1,0,
              0,0,1,0,-1,
              0,0,0,1,-1),
            nrow=3,byrow=TRUE)
rownames(X) <- c('BC','BD','CD')
colnames(X) <- c('b0','b1','b2','b3','b4')
p_beta <- coef(fit)[1:5]
mat <- X %*% p_beta

vcov(fit)
vcov_con <- X %*% vcov(fit)[1:5,1:5] %*% t(X); vcov_con

se_con <- sqrt(diag(vcov_con)); se_con
w <- (X %*% p_beta)/se_con; w
p_val <- 2 * pnorm(-1 * abs(w)); p_val
```

The null hypothesis for these contrasts is the difference between the two groups is zero. Given the p-values from the contrasts, I would conclude the following:

* Significance between B and C: DO NOT REJECT (there is no significant difference between these groups);
* Significance between B and D: REJECT (these two groups are significantly different); and 
* Significance between C and D: REJECT (these two groups are significantly different).

---

Le fin.