---
title: 'Homework #6'
author: "Darien Lozon"
date: "March 26, 2019"
output: html_document
---

1. Load data and place into an unmarkedFrameOccu object.

```{r}
# Set the stage #
setwd("C:/Users/dnl0009/Google Drive/SP19 Courses/WMAN633/")
library(unmarked)

# Load and look over data
y <- read.csv('Bobcat.csv')

# Format data
bob_mat <- as.matrix(y)
occu_dat <- unmarkedFrameOccu(y=bob_mat)
```

---

2. Fit an occupancy model that assumes conditional detection probability is a function of the number of people photographed at a site, and occupancy probability is a function of disturbance within 5km.

```{r}
# Read in covs
site_covs <- read.csv('psi covariates.csv')
det_covs <- read.csv('p covariates.csv')

occu_data <- unmarkedFrameOccu(y=bob_mat,
                               siteCovs=site_covs, # site covs
                               obsCovs=list(people=det_covs)) # detection covs
fit <- occu(~people~Dist_5km,occu_data)
summary(fit)
```

---

3. Interpret the effect of the number of people photographed at each site on conditional detection probability of bobcat. Verify your interpretation in R.

The p coefficient for people (-0.0609) indicates that the log odds ratio decreases by 0.061 with each additional person captured on a camera trap.

```{r}
b <- coef(fit)
z <- plogis(b[3]) # no people
o <- plogis(b[3] + b[4]) # one person
log((o/(1-o))/(z/(1-z))) # inverse log link
b[4] # Check
```

---

4. Re-create the Wald test for the effect of disturbance within 5 km. What is the null hypothesis? Do you reject or fail to reject this null hypothesis?

The null hypothesis being tested is disturbance will not influence occupancy probability of bobcats. Based on the p-value being 7.31e-7, I would reject the null hypothesis. Disturbance within 5 km of the camera trap does indeed influence bobcat occupancy probability.

```{r}
w <- -23.64/4.773; w
2*pnorm(-1 * abs(w)) # Difference from the output of the model could be due to rounding
```

---

5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over the range of observed disturbance within 5km.

```{r}
# Set up new site covariates
new_psi <- data.frame(Dist_5km = seq(from=min(site_covs$Dist_5km), # NEED the "Dist_5km ="
                          to=max(site_covs$Dist_5km),
                          length.out=100))
colnames(new_psi) <- "Dist_5km" # make it pretty
head(new_psi) # Check values

# Predict
predict <- predict(object=fit,newdata=new_psi,type='state')
head(predict) # Check values

# Plot
plot(x=new_psi$Dist_5km,y=exp(predict$Predicted),
     ylim=c(min(exp(predict$Predicted)),
            max(exp(predict$Predicted))),
     type='l', xlab='Change in Disturbance w/in 5km',ylab='Occupancy Probability',
     main='Effect of Disturbance on Bobcat Occupancy Probability')
lines(y=exp(predict$upper),x=new_psi$Dist_5km,col="red",lty=3) # Plot 95% confidence intervals
lines(y=exp(predict$lower),x=new_psi$Dist_5km,col="blue",lty=3)
legend("topright",c('Prediction','Upper CI','Lower CI'),col=c('black','red','blue'),lty=c(1,3,3))
```

---

Le fin.